HMM Model

### 2020-10-22 manuscript/hmm.Rmd
Optimized and cleaned HMM models

```{r}
devtools::load_all()
library(tidyverse)
options(mc.cores = parallel::detectCores())
rstan::rstan_options(auto_write = TRUE)

theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()
wide <- data$marker_data_wide

cache_dir <- here::here("local_temp_data")
if(!dir.exists(cache_dir)) {
  dir.create(cache_dir)
}

@@

#Temporary hack
registerS3method("stan_llh", class = "rate_hmm", method = stan_llh.rate_hmm, envir = asNamespace("brms"))
```

```{r}
serie_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup()

@@

```{r}
data$patient_data %>% filter(!is.na(best_supportive_care_from)) %>% select(outcome)
```

```{r}
hcq_m1_hidden_state_data <- tibble(id = factor(disease_levels, levels = disease_levels)) %>% mutate( corresponding_obs = id)

hcq_m1_observed_state_data <- tibble(id = factor(disease_levels, levels = disease_levels), is_noisy = FALSE)

recode_group <- function(rate_data, group_name) {
  is_column <- paste0("is_", group_name) 
  id_column <- paste0(group_name, "_id")

hcq_m1_rate_data <- rbind(
  tibble(.from = breathing_levels, .to = "Death", rate_group = "death"),
  tibble(.from = "AA", .to = "Discharged", rate_group = "improve_one"),
  tibble(.from = breathing_levels[2 : length(breathing_levels)], .to = breathing_levels[1 : (length(breathing_levels) - 1)], rate_group = "improve_one"),
  tibble(.from = breathing_levels[1 : (length(breathing_levels) - 1)], .to = breathing_levels[2 : length(breathing_levels)], rate_group = "worsen_one")
) %>%
  mutate(.rate_id = factor(1:n()),
         .from = factor(.from, levels = disease_levels),
         .to = factor(.to, levels = disease_levels)
  ) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one")

hcq_m1_initial_states <- serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

hcq_m1_prior = brms::set_prior("normal(-2, 5)", "Intercept") +
  brms::set_prior("normal(0, 5)", "b") +
  brms::set_prior("normal(0, 2)", "sd")

hcq_m1_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group
hcq_m1 <- brmshmmdata(hcq_m1_formula, serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

#make_stancode_hmm(hcq_m1)
#xx <- brms::make_standata(hcq_m1$formula, data = make_data_hmm(hcq_m1)$brmsdata)
hcq_m1_fit <- brmhmm(hcq_m1, cache_file = paste0(cache_dir,"/hcq_m1.rds"))

summary(hcq_m1_fit$brmsfit)

```
```{r}

# counterfactual_max_days <- 90
# counterfactual_base <-  crossing(day = 0:counterfactual_max_days, patient_id = unique(serie_data$patient_id)) %>%
#   inner_join(data$patient_data, by = c("patient_id")) %>% 
#   left_join(wide, by = c("hospital_id", "patient_id", "day")) %>%
#   mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
#   mutate(.time = .time +1) %>% 
#   filter(.time >= 1) %>%
#   group_by(.serie) %>%
#   mutate(best_supportive_care = !is.na(best_supportive_care_from) & day >= best_supportive_care_from) %>%
#   ungroup()
# 
# 
# counterfactual_took_hcq <- counterfactual_base %>% mutate(took_hcq = TRUE)
# counterfactual_no_hcq <- counterfactual_base %>% mutate(took_hcq = FALSE)
# 
# 
# 
# hcq_m1_counterfactual_took_hcq <- hcq_m1
# hcq_m1_counterfactual_took_hcq$serie_data = counterfactual_took_hcq
# 
# hcq_m1_counterfactual_took_hcq_epred_rect <- posterior_epred_rect(hcq_m1_fit, newdata = hcq_m1_counterfactual_took_hcq)
# 
# hcq_m1_counterfactual_no_hcq <- hcq_m1
# hcq_m1_counterfactual_no_hcq$serie_data = counterfactual_no_hcq
# 
# hcq_m1_counterfactual_no_hcq_epred_rect <- posterior_epred_rect(hcq_m1_fit, newdata = hcq_m1_counterfactual_no_hcq)

@@

```{r}
draws <- tidybayes::tidy_draws(hcq_m1_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m1_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "supportive"
  )

hmm_hypothesis_res_list[["hcq_m1_discharged"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "supportive"
  )
```

```{r}
if(do_pp_checks) {
  hcq_m1_epred_rect <- posterior_epred_rect(hcq_m1_fit)
  hcq_m1_predicted_rect <- posterior_epred_to_predicted(hcq_m1_fit, hcq_m1_epred_rect)
  hcq_m1_predicted <- posterior_rect_to_long(hcq_m1_fit, hcq_m1_predicted_rect)
}
```

```{r}
if(do_pp_checks) {
  hcq_m1_predicted_df <- posterior_long_to_df(hcq_m1_fit$data, hcq_m1_predicted)

  all_ids <- hcq_m1_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
    distinct() %>% arrange(hospital_id) %>% pull(.serie)
  step_size <- 6
  for(step in 1:ceiling(length(all_ids) / step_size)) {
    series_id <- ((step - 1) * step_size + 1) : (step* step_size)

    posterior_state_plot(hcq_m1_predicted_df, all_ids[series_id], hcq_m1_fit$data) %>% print()

  }
}
```

```{r}
if(do_pp_checks) {
  pp_check_last_state(hcq_m1_fit, hcq_m1_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
  pp_check_transitions_direction(hcq_m1_fit, hcq_m1_predicted_rect)
  pp_check_transitions_direction(hcq_m1_fit, hcq_m1_predicted_rect, states_from = 2:6)
}
```

```{r}
hcq_m2_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group
hcq_m2 <- brmshmmdata(hcq_m2_formula, serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

hcq_m2_fit <- brmhmm(hcq_m2, cache_file = paste0(cache_dir,"/hcq_m2.rds"))

summary(hcq_m2_fit$brmsfit)
```

```{r}
draws <- tidybayes::tidy_draws(hcq_m2_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m2_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "age, sex, supportive"
  )

hmm_hypothesis_res_list[["hcq_m2_discharged"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "age, sex, supportive"
  )
```

```{r}
hcq_m2_epred_rect <- posterior_epred_rect(hcq_m2_fit)
hcq_m2_predicted_rect <- posterior_epred_to_predicted(hcq_m2_fit, hcq_m2_epred_rect)
hcq_m2_predicted <- posterior_rect_to_long(hcq_m2_fit, hcq_m2_predicted_rect)
```

```{r}
hcq_m2_predicted_df <- posterior_long_to_df(hcq_m2_fit$data, hcq_m2_predicted)

all_ids <- hcq_m2_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m2_predicted_df, all_ids[series_id], hcq_m2_fit$data) %>% print()

}
```

```{r}
pp_check_last_state(hcq_m2_fit, hcq_m2_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m2_fit, hcq_m2_predicted_rect)
pp_check_transitions_direction(hcq_m2_fit, hcq_m2_predicted_rect, states_from = 2:6)
```

```{r}
hcq_m3_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id)
hcq_m3 <- brmshmmdata(hcq_m3_formula, serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

hcq_m3_fit <- brmhmm(hcq_m3, cache_file = paste0(cache_dir,"/hcq_m3.rds"))

summary(hcq_m3_fit$brmsfit)
```

```{r}
draws <- tidybayes::tidy_draws(hcq_m3_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m3_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "age, sex, supportive, hospital"
  )

hmm_hypothesis_res_list[["hcq_m3_discharged"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "age, sex, supportive, hospital"
  )
```

```{r}
hcq_m3_epred_rect <- posterior_epred_rect(hcq_m3_fit)
hcq_m3_predicted_rect <- posterior_epred_to_predicted(hcq_m3_fit, hcq_m3_epred_rect)
hcq_m3_predicted <- posterior_rect_to_long(hcq_m3_fit, hcq_m3_predicted_rect)
```

```{r}
hcq_m3_predicted_df <- posterior_long_to_df(hcq_m3_fit$data, hcq_m3_predicted)

all_ids <- hcq_m3_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m3_predicted_df, all_ids[series_id], hcq_m3_fit$data) %>% print()

}
```

```{r}
pp_check_last_state(hcq_m3_fit, hcq_m3_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m3_fit, hcq_m3_predicted_rect)
pp_check_transitions_direction(hcq_m3_fit, hcq_m3_predicted_rect, states_from = 2:6)
```

```{r}
hcq_m4_formula <- ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + took_az * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id)
hcq_m4 <- brmshmmdata(hcq_m4_formula, serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

hcq_m4_fit <- brmhmm(hcq_m4, cache_file = paste0(cache_dir,"/hcq_m4.rds"))

summary(hcq_m4_fit$brmsfit)
```

```{r}
draws <- tidybayes::tidy_draws(hcq_m4_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m4_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "age, sex, supportive, hospital, az"
  )

hmm_hypothesis_res_list[["hcq_m4_discharged"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "age, sex, supportive, hospital, az"
  )

hmm_hypothesis_res_list[["hcq_m4_death_az"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_azTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "age, sex, supportive, hospital, hcq"
  )

hmm_hypothesis_res_list[["hcq_m4_discharged_az"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_azTRUE + draws$`b_rate_groupimprove_one:took_azTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "age, sex, supportive, hospital, hcq"
  )
```

```{r}
hcq_m4_epred_rect <- posterior_epred_rect(hcq_m4_fit)
hcq_m4_predicted_rect <- posterior_epred_to_predicted(hcq_m4_fit, hcq_m4_epred_rect)
hcq_m4_predicted <- posterior_rect_to_long(hcq_m4_fit, hcq_m4_predicted_rect)
```

```{r}
hcq_m4_predicted_df <- posterior_long_to_df(hcq_m4_fit$data, hcq_m4_predicted)

all_ids <- hcq_m4_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m4_predicted_df, all_ids[series_id], hcq_m4_fit$data) %>% print()

}
```

```{r}
pp_check_last_state(hcq_m4_fit, hcq_m4_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m4_fit, hcq_m4_predicted_rect)
pp_check_transitions_direction(hcq_m4_fit, hcq_m4_predicted_rect, states_from = 2:6)
```

## hcq_m5 -

```{r}
hcq_m5_formula <-  ~ is_improve_one + is_death + mo(death_id) + (0 + is_worsen_one | worsen_one_id) + 
    (0 + is_improve_one | improve_one_id) + best_supportive_care * rate_group + took_hcq * rate_group + age_norm * rate_group + sex * rate_group + (rate_group | hospital_id) + comorbidities_sum * rate_group #mo(comorbidities_sum) * rate_group
hcq_m5 <- brmshmmdata(hcq_m5_formula, serie_data, 
                  hcq_m1_rate_data, hcq_m1_hidden_state_data, hcq_m1_initial_states, prior = hcq_m1_prior,
                  observed_state_data = hcq_m1_observed_state_data)

hcq_m5_fit <- brmhmm(hcq_m5, cache_file = paste0(cache_dir,"/hcq_m5.rds"))

summary(hcq_m5_fit$brmsfit)
```

```{r}
draws <- tidybayes::tidy_draws(hcq_m5_fit$brmsfit)
hmm_hypothesis_res_list[["hcq_m5_death"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_reduces_death,
    adjusted = "age, sex, supportive, hospital, comorbidities_sum"
  )

hmm_hypothesis_res_list[["hcq_m5_discharged"]] <- 
  bayesian_hypothesis_res_from_draws(
    draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
    model = "HMM",
    estimand = "log(HR)",
    hypothesis = hypotheses$hcq_increases_discharged,
    adjusted = "age, sex, supportive, hospital, comorbidities_sum"
  )
```

```{r}
hcq_m5_epred_rect <- posterior_epred_rect(hcq_m5_fit)
hcq_m5_predicted_rect <- posterior_epred_to_predicted(hcq_m5_fit, hcq_m5_epred_rect)
hcq_m5_predicted <- posterior_rect_to_long(hcq_m5_fit, hcq_m5_predicted_rect)
```

```{r}
hcq_m5_predicted_df <- posterior_long_to_df(hcq_m5_fit$data, hcq_m5_predicted)

all_ids <- hcq_m5_fit$data$serie_data %>% select(hospital_id, .serie) %>% 
  distinct() %>% arrange(hospital_id) %>% pull(.serie)
step_size <- 6
for(step in 1:ceiling(length(all_ids) / step_size)) {
  series_id <- ((step - 1) * step_size + 1) : (step* step_size)

  posterior_state_plot(hcq_m5_predicted_df, all_ids[series_id], hcq_m5_fit$data) %>% print()

}
```

```{r}
pp_check_last_state(hcq_m5_fit, hcq_m5_predicted_rect) #TODO separately terminal and non-terminal states. Or "state at time t" - this should be coherent
pp_check_transitions_direction(hcq_m5_fit, hcq_m5_predicted_rect)
pp_check_transitions_direction(hcq_m5_fit, hcq_m5_predicted_rect, states_from = 2:6)
```

## hcq_m6 - D-dimer

```{r}
hcq_m6_serie_data <- serie_data %>% 
  group_by(.serie) %>%
  filter(any(!is.na(d_dimer))) %>%
  ungroup() %>%
  compute_marker_peak(column_name = "d_dimer", new_column_name = "peak_d_dimer", initial_value = 0) %>%
  mutate(log_peak_d_dimer = log(peak_d_dimer + 1) - 4)

hcq_m6_initial_states <- hcq_m6_serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

>>>>

```{r setup}
devtools::load_all()
library(tidyverse)
options(mc.cores = parallel::detectCores())
rstan::rstan_options(auto_write = TRUE)
options(brms.backend = "cmdstanr")

theme_set(cowplot::theme_cowplot())
data <- read_data_for_analysis()

if(length(unique(data$patient_data$patient_id)) != length(data$patient_data$patient_id)) {
  stop("FAil")
}
# warning("Subsetting data")
# set.seed(4568224)
# patients_to_use <- sample(data$patient_data$patient_id, size = 70)
# data <- filter_complete_data(data, patient_id %in% patients_to_use)

wide <- data$marker_data_wide

cache_dir <- here::here("local_temp_data", "hmm")
if(!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

@@

#Temporary hack
registerS3method("stan_log_lik", class = "rate_hmm", method = stan_log_lik.rate_hmm, envir = asNamespace("brms"))
```

```{r}
serie_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup() %>%
  #Force explicit dummy coding for the models
  mutate(across(tidyselect::starts_with("took_"), as.integer),
         best_supportive_care = as.integer(best_supportive_care))
  #mutate(best_supportive_care = as.integer(best_supportive_care))

serie_data_28 <- wide %>%
  inner_join(crossing(data$patient_data, day = 0:28), by = c("hospital_id", "patient_id", "day")) %>%
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup()

# Extend terminal states for serie_data_28
serie_data_28 <- serie_data_28 %>%
  group_by(.serie) %>%
  mutate(last_day = max(day[!is.na(.observed)]),
         last_state = .observed[day == last_day],
         .observed = if_else(last_state %in% c("Discharged", "Death") & is.na(.observed) & day > last_day,
                             last_state,
                             .observed)) %>%
  ungroup()

base_observed_state_data <- tibble(id = factor(disease_s_levels, levels = disease_s_levels), is_noisy = FALSE)

simple_hidden_state_data <- tibble(id = factor(disease_s_levels, levels = disease_s_levels)) %>% mutate( corresponding_obs = id)

simple_initial_states <- serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

default_control <- list(adapt_delta = 0.9)
```

```{r}
#data$patient_data %>% filter(!is.na(best_supportive_care_from)) %>% select(outcome)
```

```{r}
recode_group <- function(rate_data, group_name) {
  is_column <- paste0("is_", group_name) 
  id_column <- paste0(group_name, "_id")

@@

# Treatments - effects on rate groups

## Only treatments

```{r trt_group_m1}
trt_group_m1_hidden_state_data <- simple_hidden_state_data

trt_group_m1_observed_state_data <- base_observed_state_data

trt_group_m1_rate_data <- rbind(
  tibble(.from = c("Oxygen", "Ventilated"), .to = "Death", rate_group = "death"),
  tibble(.from = "AA", .to = "Discharged", rate_group = "improve_one"),
  tibble(.from = breathing_s_levels[2 : length(breathing_s_levels)], .to = breathing_s_levels[1 : (length(breathing_s_levels) - 1)], rate_group = "improve_one"),
  tibble(.from = breathing_s_levels[1 : (length(breathing_s_levels) - 1)], .to = breathing_s_levels[2 : length(breathing_s_levels)], rate_group = "worsen_one")
) %>%
  mutate(.rate_id = factor(paste0(.from, "_", .to), levels = paste0(.from, "_", .to)),
         .from = factor(.from, levels = disease_s_levels),
         .to = factor(.to, levels = disease_s_levels)
  ) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one") %>%
  mutate(rate_group = fct_relevel(rate_group, "improve_one"))

trt_group_m1_initial_states <- serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

trt_group_m1_prior = #brms::set_prior("normal(-2, 5)", "Intercept") +
  brms::set_prior("normal(-2, 5)", "b") +
  brms::set_prior("normal(0, 2)", "sd")

trt_group_m1_formula <- ~ 0 + .rate_id +  (0 + best_supportive_care + took_hcq + took_az + took_favipiravir + took_convalescent_plasma || rate_group)

trt_group_m1 <- brmshmmdata(trt_group_m1_formula, serie_data, 
                  trt_group_m1_rate_data, simple_hidden_state_data, simple_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

trt_group_m1_fit <- brmhmm(trt_group_m1, cache_file = paste0(cache_dir,"/trt_group_m1.rds"), control = default_control, init = 0.1)

summary(trt_group_m1_fit$brmsfit)
```

```{r}

# counterfactual_max_days <- 90
# counterfactual_base <-  crossing(day = 0:counterfactual_max_days, patient_id = unique(serie_data$patient_id)) %>%
#   inner_join(data$patient_data, by = c("patient_id")) %>% 
#   left_join(wide, by = c("hospital_id", "patient_id", "day")) %>%
#   mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
#   mutate(.time = .time +1) %>% 
#   filter(.time >= 1) %>%
#   group_by(.serie) %>%
#   mutate(best_supportive_care = !is.na(best_supportive_care_from) & day >= best_supportive_care_from) %>%
#   ungroup()
# 
# 
# counterfactual_took_hcq <- counterfactual_base %>% mutate(took_hcq = TRUE)
# counterfactual_no_hcq <- counterfactual_base %>% mutate(took_hcq = FALSE)
# 
# 
# 
# trt_group_m1_counterfactual_took_hcq <- trt_group_m1
# trt_group_m1_counterfactual_took_hcq$serie_data = counterfactual_took_hcq
# 
# trt_group_m1_counterfactual_took_hcq_epred_rect <- posterior_epred_rect(trt_group_m1_fit, newdata = trt_group_m1_counterfactual_took_hcq)
# 
# trt_group_m1_counterfactual_no_hcq <- trt_group_m1
# trt_group_m1_counterfactual_no_hcq$serie_data = counterfactual_no_hcq
# 
# trt_group_m1_counterfactual_no_hcq_epred_rect <- posterior_epred_rect(trt_group_m1_fit, newdata = trt_group_m1_counterfactual_no_hcq)

@@

```{r}
# draws <- tidybayes::tidy_draws(trt_group_m1_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_group_m1_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "supportive"
#   )
# 
# hmm_hypothesis_res_list[["trt_group_m1_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "supportive"
#   )
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_group_m1_fit)
}
```

## Treatments + age + sex

```{r trt_group_m2}
trt_group_m2_formula <- update.formula(trt_group_m1_formula,  ~ . +  (age_norm + sex || rate_group))
trt_group_m2 <- brmshmmdata(trt_group_m2_formula, serie_data, 
                  trt_group_m1_rate_data, trt_group_m1_hidden_state_data, trt_group_m1_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m2_fit <- brmhmm(trt_group_m2, cache_file = paste0(cache_dir,"/trt_group_m2.rds"), init = 0.1)

summary(trt_group_m2_fit$brmsfit)
```

```{r}
# draws <- tidybayes::tidy_draws(trt_group_m2_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_group_m2_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive"
#   )
# 
# hmm_hypothesis_res_list[["trt_group_m2_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive"
#   )
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_group_m2_fit)
}
```

## Treatments + age + sex + hospital

```{r trt_group_m3}
trt_group_m3_formula <- update.formula(trt_group_m2_formula,  ~  . + (0 + .rate_id | hospital_id))
trt_group_m3 <- brmshmmdata(trt_group_m3_formula, serie_data, 
                  trt_group_m1_rate_data, trt_group_m1_hidden_state_data, trt_group_m1_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m3_fit <- brmhmm(trt_group_m3, cache_file = paste0(cache_dir,"/trt_group_m3.rds"), init = 0.1)

summary(trt_group_m3_fit$brmsfit)
```

```{r}
# draws <- tidybayes::tidy_draws(trt_group_m3_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_group_m3_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive, hospital"
#   )
# 
# hmm_hypothesis_res_list[["trt_group_m3_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive, hospital"
#   )
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_group_m3_fit)
}
```

## Treatment + age + sex + hospital + comorbidities

```{r trt_group_m4}
trt_group_m4_formula <-  update.formula(trt_group_m3_formula ~ . + (comorbidities_sum || rate_group)) trt_group_m4 <- brmshmmdata(trt_group_m4_formula, serie_data, 
                  trt_group_m1_rate_data, trt_group_m1_hidden_state_data, trt_group_m1_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m4_fit <- brmhmm(trt_group_m4, cache_file = paste0(cache_dir,"/trt_group_m4.rds"), init = 0.1)

summary(trt_group_m4_fit$brmsfit)
```

```{r}
# draws <- tidybayes::tidy_draws(trt_group_m4_fit$brmsfit)
# hmm_hypothesis_res_list[["trt_group_m4_death"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_reduces_death,
#     adjusted = "age, sex, supportive, hospital, comorbidities_sum"
#   )
# 
# hmm_hypothesis_res_list[["trt_group_m4_discharged"]] <- 
#   bayesian_hypothesis_res_from_draws(
#     draws = draws$b_took_hcqTRUE + draws$`b_rate_groupimprove_one:took_hcqTRUE`,
#     model = "HMM",
#     estimand = "log(HR)",
#     hypothesis = hypotheses$hcq_increases_discharged,
#     adjusted = "age, sex, supportive, hospital, comorbidities_sum"
#   )
```

```{r}
if(do_pp_checks) {
  do_common_pp_checks(trt_group_m4_fit)
}
```

# Treatments - effect on individual rates

## Only treatments

```{r trt_rate_m1}
trt_rate_m1_hidden_state_data <- simple_hidden_state_data

trt_rate_m1_observed_state_data <- base_observed_state_data

trt_rate_m1_rate_data <- trt_group_m1_rate_data

trt_rate_m1_initial_states <- trt_group_m1_initial_states

trt_rate_m1_prior = trt_group_m1_prior

trt_rate_m1_formula <- ~ 0 + .rate_id +  (0 + best_supportive_care + took_hcq + took_az + took_favipiravir + took_convalescent_plasma || .rate_id)

trt_rate_m1 <- brmshmmdata(trt_rate_m1_formula, serie_data, 
                  trt_rate_m1_rate_data, simple_hidden_state_data, simple_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = base_observed_state_data)

trt_rate_m1_fit <- brmhmm(trt_rate_m1, cache_file = paste0(cache_dir,"/trt_rate_m1.rds"), control = default_control, init = 0.1)

summary(trt_rate_m1_fit$brmsfit)
```

```{r}

# counterfactual_max_days <- 90
# counterfactual_base <-  crossing(day = 0:counterfactual_max_days, patient_id = unique(serie_data$patient_id)) %>%
#   inner_join(data$patient_data, by = c("patient_id")) %>% 
#   left_join(wide, by = c("hospital_id", "patient_id", "day")) %>%
#   mutate(.time = day, .serie = patient_id, .observed = breathing) %>%
#   mutate(.time = .time +1) %>% 
#   filter(.time >= 1) %>%
#   group_by(.serie) %>%
#   mutate(best_supportive_care = !is.na(best_supportive_care_from) & day >= best_supportive_care_from) %>%
#   ungroup()
# 
# 
# counterfactual_took_hcq <- counterfactual_base %>% mutate(took_hcq = TRUE)
# counterfactual_no_hcq <- counterfactual_base %>% mutate(took_hcq = FALSE)
# 
# 
# 
# trt_rate_m1_counterfactual_took_hcq <- trt_rate_m1
# trt_rate_m1_counterfactual_took_hcq$serie_data = counterfactual_took_hcq
# 
# trt_rate_m1_counterfactual_took_hcq_epred_rect <- posterior_epred_rect(trt_rate_m1_fit, newdata = trt_rate_m1_counterfactual_took_hcq)
# 
# trt_rate_m1_counterfactual_no_hcq <- trt_rate_m1
# trt_rate_m1_counterfactual_no_hcq$serie_data = counterfactual_no_hcq
# 
# trt_rate_m1_counterfactual_no_hcq_epred_rect <- posterior_epred_rect(trt_rate_m1_fit, newdata = trt_rate_m1_counterfactual_no_hcq)
